---
title: "Overlap of Appendix and Mice"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    includes:
      in_header: ../include/in_header.html
      before_body: ../include/before_body.html
      after_body: ../include/after_body.html
runtime: shiny_prerendered
---
<!-- # (c)  Juozas GordeviÄius -->

```{r, include=FALSE}
source("../code/common.R")

# gene mapping
gmap <- fread("../etc/padlock_genes_human2mouse.csv") %>%
  select(V1, V2)
gmap[1, ] %>% as.character %>% setnames(gmap, .)
gmap <- gmap[-1,]

gmap[, length(unique(MGI_mm10))]
gmap[, length(unique(HGNC_hg19))]

# Load mice DSS fits
dss <- fread(Sys.glob("../Mice_DSS_Padlock/www/m6_*.csv"))

# Load CP fits
cp <- fread(Sys.glob("../Mice_CecalPatch_Padlock/www/m6_*.csv"))


myFisherTest <- function(t) {
  tryCatch({
    f <- fisher.test(t)
    data.table(estimate = f$estimate, p.value = f$p.value)
  }, error = function(e) {
    message(e)
    data.table(estimate = NA, P = NA)
  })
}


# genomic annotations
p_load("annotatr")
foo <- function() {
  require(annotatr)
  genome <- "mm10"
  annots <- c('mm10_cpgs', 'mm10_basicgenes', 'mm10_genes_intergenic')
  annotations_mm10 <- build_annotations(genome = genome, annotations = annots)

  genome <- "hg19"
  annots <- c('hg19_cpgs', 'hg19_basicgenes', 'hg19_genes_intergenic')
  annotations_hg19 <- build_annotations(genome = genome, annotations = annots)
  return(list(mm10=annotations_mm10, hg19=annotations_hg19))  
}

g(annotations_mm10, annotations_hg19) %=% cache(
  foo = foo, fname = "annotations.RDS", verbose=TRUE
  )

# readers
readAppendix = function(prefix) { 
  fread(glue("../Appendix_PDvsControls_Padlock/www/{prefix}_Appendix_PDvsControls_Padlock.csv")) %>% 
  .[, list(ID, Gene, Chr, SNP, logFC, P.Value, adj.P.Val)]
}
readMice = function(prefix, contrast) {
  require(glue)
  x <- fread(glue("../Mice_DSS_Padlock/www/{prefix}_Mice_DSS_Padlock.csv"))
  x[, list(ID, Chr, SNP, Gene, logFC = get(glue("C.{contrast}")), 
    P.Value = get(glue("P.{contrast}")),
    adj.P.Val = p.adjust(get(glue("P.{contrast}")), "fdr"))] 
}
readCecalPatch <- function(prefix) {
  require(glue)
  x <- fread(glue("../Mice_CecalPatch_Padlock/www/{prefix}_Mice_CecalPatch_Padlock.csv")) %>%
    .[, list(ID, Gene, Chr, SNP, logFC, P.Value, adj.P.Val)]
}


# Enrichment of genes with significant loci
# dt must have Gene and Significant fields
computeORs <- function(dt) {
  dt <- 
    foreach (gene = unique(dt$Gene), .combine = rbind) %dopar% {
      require(data.table)
      tryCatch({
        t <- dt[, table(Gene == gene, Significant == TRUE)]
        f <- fisher.test(t)
        data.table(OR = f$estimate, P = f$p.value, Gene = gene)    
      }, error = function(e) {
        message(e)
        data.table(OR = NA, P = NA, Gene = gene)
      })
    }
  dt
}


CLUSTER <<- makeMyCluster(nNodes = 10)
```


# Overlap of Mice DSS contrasts

```{r}
prefix <- "m6"
myContrasts <- c(
  "GTwt_DSS - GTwt_Water",
  "GTtg_DSS - GTwt_Water",
  "GTtg_DSS - GTtg_Water",
  "GTtg_Water - GTwt_Water",
  "(GTwt_DSS - GTwt_Water) - (GTtg_DSS - GTtg_Water)")

data <- foreach(contrast = myContrasts) %do% {
  readMice(prefix, contrast)
}
names(data) <- myContrasts


# Compare loci in all pairwise contrasts
qT <- 0.05
n <- length(myContrasts)
dt <- 
  foreach(i = 1:(n-1), .combine=rbind) %:% 
  foreach(j = (i+1):n, .combine=rbind) %do% {
    if (i == j) break;
    contA <- myContrasts[i]
    contB <- myContrasts[j]
    stopifnot(all(c("ID", "Gene", "Chr", "SNP") %in% colnames(data[[contA]])))
    stopifnot(all(c("ID", "Gene", "Chr", "SNP") %in% colnames(data[[contB]])))
    dt <- merge(data[[contA]], data[[contB]], by = c("ID", "Gene", "Chr", "SNP"))  
    dt[, table(adj.P.Val.x < qT, adj.P.Val.y < qT)] %>%
      myFisherTest %>%
      .[, ContrastA := contA] %>%
      .[, ContrastB := contB]
  }
dt[, Star := gtools::stars.pval(p.value)]
```

Test overlap of significant loci using Fisher's test. Threshold q at 0.05.

```{r, include=TRUE}
dt[, list(ContrastA, ContrastB, estimate, p.value, Star)] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

```{r}

oddsRatios <- foreach(contrast = myContrasts) %do% {
  computeORs( data[[contrast]][, list(Gene, Significant = adj.P.Val < qT)])
}
names(oddsRatios) <- myContrasts

dt <- 
  foreach(i = 1:(n-1), .combine=rbind) %:% 
  foreach(j = (i+1):n, .combine=rbind) %do% {
    if (i == j) break;
    contA <- myContrasts[i]
    contB <- myContrasts[j]

    merge(oddsRatios[[contA]], oddsRatios[[contB]], by = "Gene") %>%
      .[OR.x > 0 & OR.y > 0] %>%
      lm( log(OR.x) ~ 0 + log(OR.y), data = .) %>% 
      summary %>%
      broom::tidy() %>%
      dplyr::select(estimate, p.value) %>%
      setDT %>%
      .[, ContrastA := contA] %>%
      .[, ContrastB := contB]
  }
dt[, Q := p.adjust(p.value, "fdr")]
dt[, Star := gtools::stars.pval(p.value)]
```

Threshold at q < 0.05 and count enrichment of gene with DMCs (OR). Test the correlation of odds ratios.

```{r, include=TRUE}
dt[, list(ContrastA, ContrastB, estimate, p.value, Q, Star)] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# Overlap of DSS and CecalPatch

```{r}
Mcp <- readCecalPatch(prefix)

qT <- 0.05
n <- length(myContrasts)
dt <- 
  foreach(i = 1:(n), .combine=rbind) %do% {
    contA <- myContrasts[i]
    dt <- merge(Mcp, data[[contA]], by = c("ID", "Gene", "Chr", "SNP"))  
    dt[, table(adj.P.Val.x < qT, adj.P.Val.y < qT)] %>%
      myFisherTest %>%
      .[, Contrast := contA] 
  }
dt[, Star := gtools::stars.pval(p.value)]

```

Test overlap of significant loci using Fisher's test. Threshold q at 0.05.

```{r, include=TRUE}
dt[, list(Contrast, estimate, p.value, Star)] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


```{r}
McpORs <- computeORs( Mcp[, list(Gene, Significant = adj.P.Val < qT)])
dt <- 
  foreach(i = 1:n, .combine=rbind) %do% {
    contA <- myContrasts[i]

    merge(McpORs, oddsRatios[[contA]], by = "Gene") %>%
      .[OR.x > 0 & OR.y > 0] %>%
      lm( log(OR.x) ~ 0 + log(OR.y), data = .) %>% 
      summary %>%
      broom::tidy() %>%
      dplyr::select(estimate, p.value) %>%
      setDT %>%
      .[, Contrast := contA]
  }
dt[, Q := p.adjust(p.value, "fdr")]
dt[, Star := gtools::stars.pval(p.value)]

fwrite(dt, file = "ORs_CecalPatch_DSS.csv")
```


Threshold at q < 0.05 and count enrichment of gene with DMCs (OR). Test the correlation of odds ratios.

```{r, include=TRUE}
dt[, list(Contrast, estimate, p.value, Q, Star)] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```


# Enrichment of pathways


```{r}

# See if any of the autophagy pathways is enriched
getGOGenes <- function(term) {
  require(glue)
  url <- "http://www.informatics.jax.org/go/report.txt?goID={term}&results=1000&startIndex=0&sort=term&dir="
  x <- fread(glue(url))
  x <- x[,2, with=FALSE]
  setnames(x, "Gene")
  unique(x$Gene)
}


pathways <- list(
  list(term="GO:0061912", name="Selective autophagy"),
  list(term="GO:0016236", name="Macroautophagy"),
  list(term="GO:0005764", name="Lysosomes"),
  list(term="GO:0061684", name="Chaperone-mediated autophagy"),
  list(term="GO:0006914", name="Autophagy")
  # list(term="GO:0016237", name="Microautophagy"),
  # list(term="GO:0000423", name="Mitophagy")
  )

pathways <- foreach(p = pathways) %do% {
  p$genes <- getGOGenes(p$term)
  p
}
lapply(pathways, `[[`, 3) %>% sapply(length)

# How did we compute enrichment in figure 1?
contrasts <- colnames(dss)
contrasts <- contrasts[grepl("P\\.", contrasts)]
contrasts <- contrasts[1:5]
pd <- 
  foreach(p = pathways, .combine = rbind) %:%
  foreach(c = contrasts, .combine = rbind) %do%
  {
    genes <- p$genes
    dt <- dss[, list(Gene, P=p.adjust(get(c), "fdr"))]
    res1 <- dt[, table( P < 0.05, Gene %in% genes)] %>% 
            fisher.test %>%
            broom::tidy() %>%
            dplyr::select(estimate, p.value) %>%
            setDT %>%
            .[, Term := p$term] %>%
            .[, Name := p$name] %>%
            .[, Contrast := c] %>%
            .[, Type := "Target"]
    res1
  }

pd2 <- foreach(p = pathways, .combine = rbind) %do% {
    genes <- p$genes
    dt <- cp[, list(Gene, P=adj.P.Val)]
    res1 <- dt[, table( P < 0.05, Gene %in% genes)] %>% 
            fisher.test %>%
            broom::tidy() %>%
            dplyr::select(estimate, p.value) %>%
            setDT %>%
            .[, Term := p$term] %>%
            .[, Name := p$name] %>%
            .[, Contrast := "Cecal Patch"] %>%
            .[, Type := "Target"]
    res1
  }

pd <- rbind(pd, pd2)
pd[, Q := p.adjust(p.value, "fdr")]
```

```{r, include=TRUE}
pd[order(p.value)]
```

# Overlap of Appendix and Mice datasets


```{r}
# read appendix
apx <- readAppendix(prefix)

# compute odds ratios 
apxORs <- computeORs(apx[, list(Gene, Significant = adj.P.Val < qT)])
apxORs <- merge(apxORs, gmap, by.x = "Gene", by.y = "HGNC_hg19")
apxORs <- apxORs[, list(OR, P, Gene = MGI_mm10)]
apxORs <- apxORs[!duplicated(Gene)]

# Correlate to mice DSS
dt <- 
  foreach(i = 1:n, .combine=rbind) %do% {
    contA <- myContrasts[i]
    merge(apxORs, oddsRatios[[contA]], by = "Gene") %>%
      .[OR.x > 0 & OR.y > 0] %>%
      lm( log(OR.x) ~ 0 + log(OR.y), data = .) %>% 
      summary %>%
      broom::tidy() %>%
      dplyr::select(estimate, p.value) %>%
      setDT %>%
      .[, Contrast := contA]
  }

# Corerlate to mice CP
dt2 <- 
  merge(apxORs, McpORs, by = "Gene") %>%
    .[OR.x > 0 & OR.y > 0] %>%
    lm( log(OR.x) ~ 0 + log(OR.y), data = .) %>% 
    summary %>%
    broom::tidy() %>%
    dplyr::select(estimate, p.value) %>%
    setDT %>%
    .[, Contrast := "Cecal Patch"] %>%
    .[]


dt <- rbind(dt, dt2)
rm(dt2)
dt[, Q := p.adjust(p.value, "fdr")]
dt[, Star := gtools::stars.pval(p.value)]

fwrite(dt, file = "ORs_Appendix_Mice.csv")
```

Threshold at q < 0.05 and count enrichment of gene with DMCs (OR). Test the correlation of odds ratios.

```{r, include=TRUE}
dt[, list(Contrast, estimate, p.value, Q, Star)] %>%
  kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover"))
```

