---
title: "Aging and Disease Interaction"
output:
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    includes:
      in_header: ../include/in_header.html
      before_body: ../include/before_body.html
      after_body: ../include/after_body.html
runtime: shiny_prerendered
---
<!-- # (c)  Juozas GordeviÄius -->



```{r, include=FALSE, echo=FALSE}
source("../code/common.R", chdir=TRUE)
p_load(broom)

computeORs <- function(dt) {
  dt <- 
    foreach (gene = unique(dt$Gene), .combine = rbind) %dopar% {
      require(data.table)
      tryCatch({
        t <- dt[, table(Gene == gene, Significant == TRUE)]
        f <- fisher.test(t)
        data.table(OR = f$estimate, P = f$p.value, Gene = gene)    
      }, error = function(e) {
        message(e)
        data.table(OR = NA, P = NA, Gene = gene)
      })
    }
  dt
}


CLUSTER <<- makeMyCluster(nNodes = 10)


# Load Age*Diagnosis fits
apxAg <- fread("../Appendix_AgeAcceleration_Padlock/www/m6_Appendix_AgeAcceleration_Padlock.csv")
apxPD <- fread("../Appendix_PDvsControls_Padlock/www/m6_Appendix_PDvsControls_Padlock.csv")
pfcAg <- fread("../Brain_AgeAcceleration_Padlock/www/m6_Brain_AgeAcceleration_Padlock.csv")
pfcPD <- fread("../Brain_PFCRep_Padlock_withGLU/www/m6_Brain_PFCRep_Padlock_withGLU.csv")
```

# Enrichment of pathways


```{r}
# Last chance, enrichment of pathways
getGoGenes <- function(term) {
  url <- glue("http://golr-aux.geneontology.io/solr/select?defType=edismax&qt=standard&indent=on&wt=csv&rows=100000&start=0&fl=bioentity_name,bioentity,bioentity_label&facet=true&facet.mincount=1&facet.sort=count&json.nl=arrarr&facet.limit=25&hl=true&hl.simple.pre=%3Cem%20class=%22hilite%22%3E&hl.snippets=1000&csv.encapsulator=&csv.separator=%09&csv.header=false&csv.mv.separator=%7C&fq=regulates_closure:%22{term}%22&fq=document_category:%22bioentity%22&fq=taxon_subset_closure_label:%22Homo%20sapiens%22&facet.field=source&facet.field=taxon_subset_closure_label&facet.field=type&facet.field=panther_family_label&facet.field=annotation_class_list_label&facet.field=regulates_closure_label&q=*:*")
  genes <- fread(url, header = FALSE)
  return(genes$V3)
}

pathways <- list(
  list(term="GO:0061912", name="Selective autophagy"),
  list(term="GO:0016236", name="Macroautophagy"),
  list(term="GO:0005764", name="Lysosomes"),
  list(term="GO:0061684", name="Chaperone-mediated autophagy"),
  list(term="GO:0006914", name="Autophagy")
  # list(term="GO:0016237", name="Microautophagy"),
  # list(term="GO:0000423", name="Mitophagy")
  )

pathways <- foreach(p = pathways) %do% {
  p$genes <- getGoGenes(p$term)
  p
}

lapply(pathways, `[[`, 3) %>% sapply(length)
```

## In Appendix

```{r}
# Which pathway has more than expected significant DMCs?

pd <-
  apxAg[, list(ID, Gene, C = C.DiagnosisControl.Age, 
              P=P.DiagnosisControl.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))]
# pd <-
#   apxAg[, list(ID, Gene, C = C.DiagnosisPD.LBD.Age, 
#               P=P.DiagnosisPD.LBD.Age, 
#               Q = p.adjust(P.DiagnosisPD.LBD.Age, "fdr"))]

res <- foreach( p = pathways, .combine = rbind) %do% {
  message(p$name)
  pd[, table(Gene %in% p$genes, Q < 0.05)] %>%
  fisher.test %>%
  broom::tidy() %>%
  select(estimate, p.value) %>%
  setDT %>% 
  .[, Pathway := p$name]
}
```

```{r, include=TRUE}
res
```

## In PFC


```{r}
pd <- pfcAg[, list(ID, Gene, 
              C = C.GroupCTRL.Age, 
              P=P.GroupCTRL.Age, 
              Q = p.adjust(P.GroupCTRL.Age, "fdr"))]

res <- foreach( p = pathways, .combine = rbind) %do% {
  message(p$name)
  pd[, table(Gene %in% p$genes, Q < 0.05)] %>%
  fisher.test %>%
  broom::tidy() %>%
  select(estimate, p.value) %>%
  setDT %>% 
  .[, Pathway := p$name]
}
```

```{r, include=TRUE}
res
```


# Aging hallmarks

```{r}
p_load("RobustRankAggreg")

eApx <- 
  apxAg[, list(ID, Gene, C = C.DiagnosisControl.Age, 
              P=P.DiagnosisControl.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))] %>%
  .[, list(Gene, Significant = Q < 0.05)] %>%
  computeORs

ePFC <- 
  pfcAg[, list(ID, Gene, C = C.GroupCTRL.Age, 
              P=P.GroupCTRL.Age, 
              Q = p.adjust(P.GroupCTRL.Age, "fdr"))] %>%
  .[, list(Gene, Significant = Q < 0.05)] %>%
  computeORs

hallmarks <- aggregateRanks(
  list(
    ePFC[order(P), Gene],
    eApx[order(P), Gene]
  )
)

X <- head(hallmarks, 10) %>% 
  setDT %>% 
  setnames(c("Gene", "Score"))
pd <- merge(X, ePFC[, list(Gene, EPFC=-log10(P))], by = "Gene") %>%
  merge(., eApx[, list(Gene, EAPX=-log10(P))], by = "Gene") %>%
  .[order(Score)]
pd[, Width := 1/(Score / sum(Score))]
pd[, Width := Width / sum(Width)]
pd[, sum(Width)]
pd[, Start := cumsum(c(0, Width[-.N]))]
pd[, End := cumsum(Width)]
```

```{r, include=TRUE}
pd
```

# Correlation of aging and disease effects

## In appendix

Using q < 0.02 as significance threshold. 

```{r}
pd <- copy(apxAg)
pd[, CTRL := p.adjust(P.DiagnosisControl.Age, "fdr") < 0.2]
pd[, PD := p.adjust(P.DiagnosisPD.LBD.Age, "fdr") < 0.2]
pd[, DiffA := p.adjust(`P.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, "fdr") < 0.2]
```

Overlap of aging in controls and in PDs

```{r, include=TRUE}
pd[, table(CTRL, PD)] %>% tee %>% fisher.test()
```

Overlap of healthy aging and differentially aging loci:

```{r, include=TRUE}
pd[, table(CTRL, DiffA)] %>% tee %>% fisher.test
```

Overlap of PD aging and differentially aging loci:
```{r, include=TRUE}
pd[, table(PD, DiffA)] %>% tee %>% fisher.test
```

> Most of the loci did not age in healthy but started showing aging effects in PD


Overlap of disease and aging loci

```{r, echo = TRUE, include=TRUE}
merge(
  apxAg[, list(ID, C = C.DiagnosisControl.Age, 
              P=P.DiagnosisControl.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(Q < 0.2, adj.P.Val < 0.2)] %>%
  tee %>% fisher.test()


merge(
  apxAg[, list(ID, C = C.DiagnosisPD.LBD.Age, 
              P=P.DiagnosisPD.LBD.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(Q < 0.2, adj.P.Val < 0.2)] %>%
  tee %>% fisher.test()


merge(
  apxAg[, list(ID, C = `C.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, 
              P=`P.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, 
              Q = p.adjust(`P.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(Q < 0.2, adj.P.Val < 0.2)] %>%
  tee %>% fisher.test()

```

Disease effect size correlation with aging effect size

```{r, echo = TRUE, include=TRUE}
merge(
  apxAg[, list(ID, C = C.DiagnosisControl.Age, 
              P=P.DiagnosisControl.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  # .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test(abs(C), abs(logFC), use='p')]

merge(
  apxAg[, list(ID, C = C.DiagnosisPD.LBD.Age, 
              P=P.DiagnosisPD.LBD.Age, 
              Q = p.adjust(P.DiagnosisControl.Age, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  # .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test(abs(C), abs(logFC), use='p')]

merge(
  apxAg[, list(ID, C = `C.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, 
              P=`P.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, 
              Q = p.adjust(`P.DiagnosisControl.Age - DiagnosisPD.LBD.Age`, "fdr"))],
  apxPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  # .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test(abs(C), abs(logFC), use='p')]

```

## In PFC brain

Using q < 0.02 as significance threshold. 

```{r}
pd <- copy(pfcAg)
pd[, CTRL := p.adjust(P.GroupCTRL.Age, "fdr") < 0.2]
pd[, PD := p.adjust(P.GroupPD.Age, "fdr") < 0.2]
pd[, DiffA := p.adjust(`P.GroupCTRL.Age - GroupPD.Age`, "fdr") < 0.2]
```

Overlap of aging in controls and in PDs

```{r, include=TRUE}
pd[, table(CTRL, PD)] %>% tee %>% fisher.test()
```

Overlap of healthy aging and differentially aging loci:

```{r, include=TRUE}
pd[, table(CTRL, DiffA)] %>% tee %>% fisher.test
```

Overlap of PD aging and differentially aging loci:
```{r, include=TRUE}
pd[, table(PD, DiffA)] %>% tee %>% fisher.test
```

> Most of the loci did not age in healthy but started showing aging effects in PD

Overlap of disease and aging loci

```{r, echo = TRUE, include=TRUE}
merge(
  pfcAg[, list(ID, C = C.GroupCTRL.Age, 
              P=P.GroupCTRL.Age, 
              Q = p.adjust(P.GroupCTRL.Age, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(P < 0.05, adj.P.Val < 0.05)] %>%
  tee %>% fisher.test()


merge(
  pfcAg[, list(ID, C = C.GroupPD.Age, 
              P=P.GroupPD.Age, 
              Q = p.adjust(P.GroupCTRL.Age, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(P < 0.05, adj.P.Val < 0.05)] %>%
  tee %>% fisher.test()


merge(
  pfcAg[, list(ID, C = `C.GroupCTRL.Age - GroupPD.Age`, 
              P=`P.GroupCTRL.Age - GroupPD.Age`, 
              Q = p.adjust(`P.GroupCTRL.Age - GroupPD.Age`, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  .[, table(P < 0.05, adj.P.Val < 0.05)] %>%
  tee %>% fisher.test()

```

Disease effect size correlation with aging effect size

```{r, echo = TRUE, include=TRUE}
merge(
  pfcAg[, list(ID, C = C.GroupCTRL.Age, 
              P=P.GroupCTRL.Age, 
              Q = p.adjust(P.GroupCTRL.Age, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  #.[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test( abs(C), abs(logFC), use='s')]

merge(
  pfcAg[, list(ID, C = C.GroupPD.Age, 
              P=P.GroupPD.Age, 
              Q = p.adjust(P.GroupPD.Age, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  # .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test(abs(C), abs(logFC), use='p')]

merge(
  pfcAg[, list(ID, C = `C.GroupCTRL.Age - GroupPD.Age`, 
              P=`P.GroupCTRL.Age - GroupPD.Age`, 
              Q = p.adjust(`P.GroupCTRL.Age - GroupPD.Age`, "fdr"))],
  pfcPD[, list(ID, logFC, P.Value, adj.P.Val)],
  by = "ID") %>%
  # .[P < 0.05 & P.Value < 0.05, ] %>%
  .[, cor.test(abs(C), abs(logFC), use='p')]

```



